---
title: "Week_6-Data Transformation"
output: html_document
Author: Gayathri Mutyala
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r echo=TRUE, message=FALSE, warning=FALSE}
library(nycflights13)
library(tidyverse)
#View(flights)

```

#Data Filtering Some time we need to filter the data as per our data analysis requiremnts so Following Filter techniques are used in R

```{r echo=TRUE, message=FALSE, warning=FALSE}
Aug31<-filter(flights, month == 8, day == 31)
Aug31
Feb<-filter(flights, month == 2)
Feb
Dec25<-filter(flights, month == 12, day == 25)
Dec25
```

# Using logical operatios to filter  the data and do data manupulations
```{r echo=TRUE, message=FALSE, warning=FALSE}
feb1to15<-filter(flights, month==2, day==1|day==15)
#filtering data from month of feb and dates between 1 to 15 th.

```


Finding the missing values in the given dataset
```{r echo=TRUE, message=FALSE, warning=FALSE}
#is.na(feb1to15)
# fasle denotes no missing data and True denotes the places of missing data
str(feb1to15)

```

#Arranging rows as per our requirement assending or desending 

```{r echo=TRUE, message=FALSE, warning=FALSE}

arrange(Aug31, desc(origin))
#Arranging the list of flights from aug 31 st in a decending order of origin

```

#Shuffling columns front and back as we want to look at the data and filtering required columns
```{r echo=TRUE, message=FALSE, warning=FALSE}
 
select(flights, flight, tailnum, origin, dest)   #filtering requeried columns 
select(Dec25, flight,carrier, tailnum, origin, dest, everything()) # add the required columns to the first few columns. everything pulls all the columns from the data set.

```


#Rename column names
```{r echo=TRUE, message=FALSE, warning=FALSE}
rename(Feb, tail_num = tailnum)

#Changed the coulmn name tail_num into tailnum

```


Adding new columns to the data set
```{r echo=TRUE, message=FALSE, warning=FALSE}
mutate(feb1to15, speed= distance / air_time * 60, traveltime=sched_arr_time+arr_delay)

#added speed column by using distance and time columns and added travel time column by adding scheduled time and departule delays
```


```{r echo=TRUE, message=FALSE, warning=FALSE}



```



